---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
####1.Load the data:

```{r, echo=TRUE}
activity <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
```

Quick look to data:
```{r, echo=TRUE}
str(activity)
```

```{r, echo=TRUE}
head(activity)
```

####2.Processing dates
```{r, echo=TRUE}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
activity$interval <- factor(activity$interval)
```

## What is mean total number of steps taken per day?
####1.Calculate the total number of steps taken per day
```{r, echo=TRUE}
su <- tapply(activity$steps, activity$date, sum, na.rm=T)
print(su)
```

####2.Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
hist(su, xlab = "sum of steps per day", main = "histogram of steps per day")
```

####3.Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean <- round(mean(su))
median <- round(median(su))

print(sprintf("The mean is %f",mean))
print(sprintf("The median is %f",median))
```

## What is the average daily activity pattern?
####1.Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
mn_int <- tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(mn_int ~ unique(activity$interval), type = "1", xlab = "5-minute interval")
```

####2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
mn_int[which.max(mn_int)]
```

## Imputing missing values
####1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
table(is.na(activity) == TRUE)
```

####2.Devise a strategy for filling in all of the missing values in the dataset.
####3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
activity2 <- activity  # creation of the dataset that will have no more NAs
for (i in 1:nrow(activity)){
    if(is.na(activity$steps[i])){
        activity2$steps[i]<- mn_int[[as.character(activity[i, "interval"])]]
    }
}
```

####4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r, echo=TRUE}
su2 <- tapply(activity2$steps, activity2$date, sum, na.rm=T)
hist(su2, xlab = "sum of steps per day", main = "histogram of steps per day")

mean <- round(mean(su2))
median <- round(median(su2))

print(sprintf("The new mean is %f",mean))
print(sprintf("The new median is %f",median))
```

## Are there differences in activity patterns between weekdays and weekends?
####1.Create a new factor variable in the dataset with two levels  - "weekday" and "weekend"" indicating whether a given date is a weekday or weekend day.
```{r, echo=TRUE}
#Creating a factor variable "day "to store the day of the week:
activity2$day <- as.factor(weekdays(activity2$date))

#Creating a logical variable "is_weekday" (weekday=TRUE, weekend = FALE) :
activity2$is_weekday <- ifelse(!(activity2$day %in% c("Saturday","Sunday")), TRUE, FALSE) 

#Calculating the average number of steps for weekdays
weekdays_data <- activity2[activity2$is_weekday,]
steps_per_interval_weekdays <- aggregate(weekdays_data$steps, by=list(interval=weekdays_data$interval), FUN=mean)

#Calculating the average number of steps for weekends
weekends_data <- activity2[!activity2$is_weekday,]
steps_per_interval_weekends <- aggregate(weekends_data$steps, by=list(interval=weekends_data$interval), FUN=mean)

#Adding columns names
colnames(steps_per_interval_weekdays) <- c("interval", "average_steps")
colnames(steps_per_interval_weekends) <- c("interval", "average_steps")

#Adding a column to indecate the day
steps_per_interval_weekdays$day <- "Weekday"
steps_per_interval_weekends$day <- "Weekend"

#Merging the two together
week_data <- rbind(steps_per_interval_weekends, steps_per_interval_weekdays)

#Converting the day variable to a factor
week_data$day <- as.factor(week_data$day)
```

####2.Make a panel plot
```{r, echo=TRUE}
library(lattice)
xyplot(average_steps ~  interval | day, data = week_data, layout = c(1,2), type ="l", ylab="Number of Steps")
```